# Processing Largier Lab Russian River Estuary Data

```{r include = FALSE}
library(rremat)
library(dplyr)
```

## Processing the bathymetry

This code reproduces the dataframe `volumes` from the text files in
extdata/bathymetry. The volume lookup table is generated from a 1ft x 1ft 
bathymetry raster with a vertical resolution as defined in 
`bathymetry/zgrid.csv`.

```{r process-lookup}
# load the bathymetry tables
volpath = system.file("extdata/bathymetry/bathytables", package = "rremat")
bathytables = dir(volpath, pattern = "*.csv$", full.names = TRUE)
bathydat = read_bathymetry_tables(bathytables)
# build the habitat table
wses = seq(2.7, 3.0, by = 0.1)
lookupvalues = setNames(expand.grid(unique(bathydat$bedelev), 
  unique(bathydat$bedelev)), c("elev", "wse"))
lookupvalues = filter(lookupvalues, elev <= wse)
volumes = evalmany(build_habitat_table, e = lookupvalues$elev, 
  w = lookupvalues$wse, method = "set", default.args = list(d = bathydat), 
  packages = "dplyr", clusters = 3)[1:9]
# add bathymetry resolution information
respath = system.file("extdata/bathymetry/bathymetryresolution.csv", 
  package = "rremat")
attr(volumes, "resolution") = read.csv(respath)
```

The stage-volume curve is generated from the volume lookup table.

```{r}
stage_volume = volumes  %>% 
  gather(type, habcount, count, count.littoral, count.limnetic, 
    count.epibenthic, count.sublimnetic, count.profundal) %>% 
  group_by(type, wse) %>% summarize(count = sum(habcount))
```

## Processing closure data

information on closures is contained in `extdata/closres/closures.csv`.
This code reproduces the data frame `closures` which provides meta data for
closures occurring during the monitoring periods.

```{r process-closures}
closurepath = system.file("extdata/closures", package = "rremat")
closures = read.csv(paste0(closurepath, "/closures.csv"), 
  stringsAsFactors = FALSE, colClasses = c(start = "Date", end = "Date", 
  code = "factor", note = "character"))
```

## Processing CTD transect meta data

CTD transect meta data is contained in `extdata/ctd/transectdatetimes.csv`.
This code reproduces the dataframe `ctdmeta` which provides meta data for the 
transect data described in the following sections.

```{r process-ctdmeta}
ctdmpath = system.file("extdata/ctd_meta", package = "rremat")
ctdm = read.csv(paste0(ctdmpath, "/transectdatetime.csv"), 
  stringsAsFactors = FALSE, colClasses = "character")
ctdmeta = data.frame(id = factor(ctdm$id), 
  start = as.POSIXct(paste(ctdm$date, ctdm$start), tz = "US/Pacific"),
  end = as.POSIXct(paste(ctdm$date, ctdm$end), tz = "US/Pacific"), 
  numcasts = as.integer(ctdm$numcasts), note = ctdm$note)
# cross-reference with closures
ctdmeta["code"] = factor(unlist(sapply(as.Date(ctdmeta$start), function(x){ 
  this = as.character(closures$code[which(x >= closures$start & x <= closures$end)])
  if(length(this) > 0)
    this
  else
    "O"  
  })))
# set time zones to UTC
attr(ctdmeta$start, "tzone") = "UTC"
attr(ctdmeta$end, "tzone") = "UTC"
```

## Processing CTD transect data

This code reproduces the dataframe `ctd` from the Matlab files in 
`extdata/ctd`.

```{r process-ctd}
ctdpath = system.file("extdata/ctd", package = "rremat")
ctd.list = Map(function(fpath) read_rrectd(fpath, quiet = TRUE)[[1]], 
  fpath = paste0(ctdpath, "/", dir(ctdpath)))
ctd = correct_dist(depth_from_elev(merge_rrectd(ctd.list)))
```

## Processing gridded CTD data

This code reproduces the dataframe `grids` from the matlab files
in `extdata/ctd_interp`. These grids may contain extrapolated data including 
regions outside of the estuary bathymetry and above the observed water-surface 
elevation. Note that the `dist` values have already been corrected as per the 
function `correct_dist` (see `matlab-src/interpctd.m`). Volume information and 
water-surface elevation is added to the gridded data 
using the volume lookup table and and CTD cast data, respectively. Extrapolated
data occurring either outside of the estuary bathymetry (`volume == 0`) or 
above the water-surface elevation (`elev > wse`) is removed.


```{r process-grids}
gridpath = system.file("extdata/ctd_interp", package = "rremat")
grid.list = Map(function(fpath) read_rrectdgrid(fpath, quiet = TRUE)[[1]],
  fpath = paste0(gridpath, "/", dir(gridpath)))
grids = merge_rrectd(grid.list)
grids = join_wse(grids, ctd) %>% mutate(wse = round(wse, 1)) %>% 
  filter(elev <= wse) %>% join_volume(volumes) %>% 
  filter(volume.total > 0)
```

## Processing water level pressure gauge data

This code reproduces the dataframe `wll` from the Matlab files in 
`extdata/wll`.

```{r process-wll}
wllpath = system.file("extdata/wll", package = "rremat")
wll.list = Map(function(fpath) read_rrewll(fpath, quiet = TRUE)[[1]], 
  fpath = paste0(wllpath, "/", dir(wllpath))) 
wll = merge_rrewll(wll.list)
wll = wll[!is.nan(wll$depth),]
```

## Supplemental Data

The following data sets are produced outside of the Largier Lab, 
but provide critical supplemental data.

### Tide data

This code reproduces the dataframe `tides` generated from files in 
`extdata/tides`. The files contain data for the 
[NOAA Point Reyes Station](http://tidesandcurrents.noaa.gov/stationhome.html?id=9415020) 
downloaded from the NOAA web API. Those files were created using the function 
`download_tides`.

```{r process-tides}
tidespath = system.file("extdata/tides", package = "rremat")
tides.list = Map(read.csv, file = paste0(tidespath, "/", dir(tidespath)),
  stringsAsFactors = FALSE)
tides = merge_tides(tides.list)
```

### Streamflow data

This code reproduces the dataframe `inflows` from the files in 
`extdata/streamflow`. The files contain streamflow data for from USGS gauges near 
[Guerneville, CA](http://waterdata.usgs.gov/nwis/inventory/?site_no=11467000&agency_cd=USGS)
(Russian River) and near 
[Cazadero, CA](http://waterdata.usgs.gov/nwis/inventory/?site_no=11467200&agency_cd=USGS) 
(Austin Creek). Those files were downloaded from the USGS 
[National Water Information System]([http://waterdata.usgs.gov/nwis/](http://waterdata.usgs.gov/nwis/)).

```{r process-streamflow}
sfpath = system.file("extdata/streamflow", package = "rremat")
sf.list = Map(read.csv, file = paste0(sfpath, "/", dir(sfpath)), 
  stringsAsFactors = FALSE)
inflows = merge_streamflow(sf.list)
```
